---
title: "Forecasting the Sweet Spot: A Time Series Analysis on Cocoa Prices"
format: pdf
editor: visual
author: Ethan Wu [1008507661], Rebecca Li [1008885949]
date: April 4, 2025
---

```{r}
#| include: false

## Load Packages ##
library(astsa)
library(tidyverse)
library(forecast)
library(tseries)
library(xts)
library(dplyr)
```

# Abstract

The present study investigates the application of time series analysis techniques to model and forecast cocoa futures prices. Commodity price forecasting is a crucial task in fields such as finance, agriculture, and trade, as price volatility can significantly impact economic decision-making. Cocoa prices are subject to numerous external factors, including weather conditions, geopolitical events, supply chain disruptions, and market speculation. The purpose of this study is to develop and evaluate predictive models for cocoa futures prices using historical price data and external variables such as climate factors. The datasets used in this study include daily cocoa futures prices from the International Cocoa Organization (ICCO) and climate data from Ghana, sourced from the National Centers for Environmental Information (NCEI). A range of statistical and machine learning models will be explored, including classical time series models (ARIMA, GARCH) and more advanced approaches (e.g. LSTM neural networks). The results of this study will help to assess the effectiveness of various forecasting models and provide insights into the key factors driving cocoa price movements. Future research could expand on this work by incorporating additional macroeconomic indicators and alternative modeling techniques to improve forecasting accuracy.

# Introduction

The ability to predict commodity prices is critical for financial markets, agricultural stakeholders, and policymakers. Cocoa, a vital ingredient in chocolate production, experiences price fluctuations due to various supply and demand factors, including climate conditions, geopolitical instability, and speculative trading. The objective of this study is to develop a robust forecasting model for cocoa futures prices by leveraging time series analysis techniques. Two key research questions arise from this study: whether time series models effectively predict cocoa futures prices based on historical data, and how external factors (such as climate variables) influence cocoa price movements. It is hypothesized that time series models will be effective in predicting cocoa futures prices based on historical data, and that relevant climate variables will have a significant effect on the movement of cocoa prices.

The data section of this report will describe the datasets used in this study and the preprocessing steps applied. Next, the literature review section will summarize and analyze existing research on time series forecasting and commodity price modelling, highlighting common approaches and limitations. The methodology section will then outline the statistical models chosen for forecasting, including the rationale for each approach. Next, the results section will present the findings of the analysis, comparing model performance using appropriate evaluation metrics. Finally, the discussion and conclusion section will interpret these findings, address main study limitations, and propose future research directions.

# Data

## Data Collection

The present study utilizes two primary datasets: the cocoa futures price data, and the Ghana climate data. The cocoa futures prices data consists of daily closing prices for cocoa futures contracts from ICCO, covering the period from March 10, 1994 to February 27, 2025. The Ghana climate data consists of daily records of temperature and precipitation from the NCEI, providing climate insights for one of the largest cocoa-producing countries.

## Data Cleaning and Preparation

To prepare these datasets for analysis, several preprocessing steps were applied. Missing values in the climate dataset were handled by interpolation where appropriate, and extreme outliers in the price dataset were examined to determine if they were due to reporting errors, or if they should remain in the dataset. Additionally, all variables were standardized to ensure comparability (across measurements and datasets).

Initial visualizations of the cocoa price data reveal strong seasonal patterns, suggesting that certain times of the year exhibit recurring price trends. A decomposition analysis further highlights the presence of long-term trends and cyclical behaviour in the data. Correlation analyses between cocoa prices and climate variables suggest that temperature and precipitation fluctuations may play a role in cocoa price movements, which will be further investigated in the modeling section.

## Descriptive Statistics and Visualizations

```{r, fig.width = 5, fig.height = 3, fig.align = 'center'}
#| echo: false

## Initialize Cocoa Data ##
cocoa_data <- read.csv("Daily Prices_ICCO.csv")
cocoa_data <- cocoa_data %>% rename(Price = ICCO.daily.price..US..tonne.)
cocoa_data$Price <- as.numeric(gsub(",", "", cocoa_data$Price))
cocoa_data$Date <- format(as.Date(cocoa_data$Date, format = "%d/%m/%Y"),
                          "%d/%m/%y")

cocoa_ts <- ts(cocoa_data$Price, frequency = 260)
plot(cocoa_ts, main = "Plot of Cocoa Price Time Series", 
     xlab = "Time", ylab = "Prices")
```

The time series plot of cocoa prices shows volatility early on, with prices spiking before sharply declining. However, the rest of the series fluctuates within a more stable range, with visible trends but no obvious seasonality. The variance of this time series appears to change over time, which indicates that models such as GARCH may be appropriate. Differencing may also be needed to achieve stationarity.

```{r, fig.width = 5, fig.height = 3, fig.align = 'center'}
#| echo: false
#| warning: false

## Stationarity ##
invisible(acf2(cocoa_ts, 50, main = "ACF of Cocoa Prices Time Series"))
adf.test(cocoa_ts)
```

The ACF plot shows a very gradual decay, with autocorrelations remaining positive and significant over a large number of lags. This indicates clear non-stationarity in the series -- the mean of this series is not constant over time, and differencing will likely be required. The PACF plot shows a sharp cutoff after lag 1, which is consistent with the presence of an autoregressive component in the data. Together, these patterns suggest that an ARIMA model may be appropriate, specifically the ARIMA(1,1,0) model.

The ADF test result ($p=0.01$) suggests that the cocoa price series is stationary, as the null hypothesis is rejected. However, this contrasts with the ACF plot, which shows a slow decay indicative of non-stationarity. This may be due to the high persistent autocorrelationsor structural shifts in the data that the ADF test does not fully capture. Despite the test result, the visual evidence supports differencing to ensure robustness in modeling.

```{r, fig.width = 5, fig.height = 3, fig.align = 'center'}
#| echo: false

## Heteroskedasticity ##
n <- length(cocoa_ts)
var(cocoa_ts[1:n/2])
var(cocoa_ts[(n/2):n])
## use a2 code to make this pretty ##
```

The variance in the first half of the cocoa price series is substantially larger (2,359,043) than in the second half (278,878.2) of the series, indicating a significant decrease in variability over time. This strong difference in variance confirms the presence of heteroskedasticity in the series, suggesting non-constant volatility and the potential need for techniques such as GARCH to properly represent the data.

```{r, fig.width = 5, fig.height = 3, fig.align = 'center'}
#| echo: false

## First Differencing ##
cocoa_diff <- diff(cocoa_ts)
plot(cocoa_diff, main = "Plot of First-Differenced Cocoa Prices Time Series",
     xlab = "Time", ylab = "Prices")
```

This plot shows that differencing has successfully removed the long-term trend, with the series now fluctuating around a stable mean. However, high volatility still remains in the early portion of the series, with large spikes in both directions. This indicates that while first differencing may have helped to address non-stationarity in the mean, the issue of heteroskedasticity still remains and may impact model performance. The explicit modeling of volatility through a GARCH model thus may be necessary.

```{r, fig.width = 5, fig.height = 3, fig.align = 'center'}
#| echo: false

## ACF and PACF of Differenced Series ##
invisible(acf2(cocoa_diff, 50, 
               main="ACF of First-Differenced Cocoa Price Time Series"))
```

The ACF and PACF plots for the first-differenced cocoa price time series show that autocorrelations drop off quickly and remain within the significance bounds after the first few lags -- this indicates that differencing has removed the non-stationarity present in the time series. Both the ACF and PACF plots show a significant spike at lag 1, suggesting that an ARIMA(1,1,0) model may be appropriate for capturing the trends and patterns present in the data.

```{r, fig.width = 5, fig.height = 3, fig.align = 'center'}
#| echo: false

## Initialize Climate Data ##
ghana_data <- read.csv("Ghana_data.csv")
ghana_data$PRCP[is.na(ghana_data$PRCP)] <- 0
ghana_data$DATE <- as.Date(ghana_data$DATE)

## Multivariate Time Series ##
ghana_ts <- xts(ghana_data[, c("TMAX", "TMIN", "TAVG", "PRCP")],
                order.by = ghana_data$DATE, ylab = c(-5, 120))
plot(ghana_ts, main = "Plot of Ghana Data Time Series",
     col = c("green", "red", "blue", "purple"), xlab = "Time",
     ylab = "Temperature", ylim = c(0, 110))
```

This time series plot shows strong and consistent seasonal patterns in the temperature variables (TMIN, TAVG, and TMAX), with temperatures peaking and dipping in regular yearly cycles from 1990 to 2024. The amplitude and shape of these cycles remain relatively stable over time, suggesting a predictable seasonal structure to the data. The precipitation series (purple) appears more irregular, with sporadic spikes and no clear seasonal trend -- this indicates greater variability. From this plot, temperature data could be useful as seasonal predictors in modeling cocoa prices.

```{r}
#| echo: false

## One Daily Average Temperature ##
ghana_data_avg <- ghana_data %>% group_by(DATE) %>%
  summarize(TAVG = mean(TAVG, na.rm = TRUE))
```

Note that the plot shows large chunks of missing data from the TMAX and TMIN variables, which may be difficult to extrapolate. Therefore, the best course of action here is to proceed with only the TAVG variable as a predictor for modeling cocoa prices. From the nature of this dataset, there are multiple observations every day from different stations -- therefore the overall average of the average temperatures from each date are taken to obtain one final average nationwide temperature for each day.

```{r, fig.width = 5, fig.height = 3, fig.align = 'center'}
#| echo: false

## Modify Ghana Data Date Format to Match Cocoa Data ##
ghana_data_avg$DATE <- format(as.Date(ghana_data_avg$DATE), "%d/%m/%y")
ghana_data_final <- ghana_data_avg[ghana_data_avg$DATE %in% cocoa_data$Date, ]

## Final Time Series ##
ghana_ts_final <- ts(ghana_data_final$TAVG, frequency = 260)
plot(ghana_ts_final, 
     main = "Plot of Average Ghana Daily Temperature Time Series",
     xlab = "Time", ylab = "Temperature")
```

This plot shows a clear and consistent seasonal pattern, with temperatures oscillating regularly throughout the year. This strong seasonality, as well as relatively stable variance and no evident long-term trend, suggests that temperature follows a stationary seasonal process. These characteristics make TAVG a suitable and stable predictor for forecasting cocoa prices.

```{r, fig.width = 5, fig.height = 3, fig.align = 'center'}
#| echo: false

## ACF and PACF for Final Time Series ##
invisible(acf2(ghana_ts_final, 50,
               main = "ACF of Average Ghana Daily Temperature Time Series"))
```

The ACF plot of the average daily temperature time series shows strong autocorrelation at early lags and a slow, gradual decay, which is characteristic of a seasonal and potentially non-stationary process. The PACF plot shows a series of spikes that taper off, suggesting the presence of both autoregressive and seasonal components. These patterns support the conclusion that temperature follows a highly persistent seasonal process.

```{r, fig.width = 5, fig.height = 3, fig.align = 'center'}
#| echo: false
#| warning: false

## ADF Test for Stationarity ##
adf.test(ghana_ts_final)
```

The ADF test result for this time series, $p=0.01$, indicates that the series is stationary. Combined with the ACF and PACF plots, this confirms that while the series is stationary, it does contain strong seasonal autocorrelation to be accounted for in modeling.

```{r, fig.width = 5, fig.height = 3, fig.align = 'center'}
#| echo: false

## Cross-Correlation ##
original_cex <- par(cex.main = 0.8)
ccf(ghana_ts_final, cocoa_ts,
    main = "Cross-Correlation Between Average Daily Temperature Series and Cocoa Price Series")
par(cex.main = original_cex)
```

The cross-correlation plot shows statistically significant negative correlations at many lags, indicating that past values of average temperature are inversely related to future cocoa prices. This suggests a lagged effect, where higher temperatures may lead to lower cocoa prices after a delay. This is potentially due to climate impacts on production or harvesting cycles.









































